---
title: "Presence of Polarzing Words in Australian Media"
description: |
  Welcome to the website. I hope you enjoy it!
site: distill::distill_website

---
## Introduction
In this project I hope to investigate the presence of the word "protest" and "riot" in Australian media. After searching the web I found an data source containing article headline, content and publish date from the Australian Broadcasting Corporation or ABC. 

## Importance
Its important to see if a news source is disproportionately using words like riot in their publications because its an indicator that their publications have some sort of bias. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#Distill
library(distill)

# Display Libraries
library(ggplot2)
library(ggthemes)
library(tidyverse)

#Verification
search()
```

```{r reading csv, include=FALSE}

data_abc <- read.csv("abcnews-date-text.csv")

#Viewing Data
print(data_abc)
```

## Presence of riot and protest from 2003 to 2021
I graphed the number of occurrences of each word during each year since 2003. This was done by grouping by year and summing the total number of occurrences per year. We can see the word protest in blue and the word riot in red.

```{r plot-1: years}

#plot 1 is in years and is displayed right here
data_abc <- read.csv("abcnews-date-text.csv")

data_abc |> 
  mutate(riot = str_detect(headline_text, "riot")) |> 
  mutate(protest = str_detect(headline_text, "protest")) |> 
  filter(protest == TRUE | riot == TRUE)  |> 
  mutate(protest = as.integer(protest), riot = as.integer(riot)) |> 
  mutate(year = substr(publish_date, 1, 4)) |> 
    group_by(year) |> 
  mutate(num_riot = sum(riot)) |> 
  mutate(num_protest = sum(protest)) |> 
  ggplot(aes(year)) +
      geom_point(aes(y = num_riot), color = "darkred") +
      geom_point(aes(y = num_protest), color = "blue") +
      theme_clean() +
      labs(title = "Presence of 'riot' and 'protest' in Australian Media Over Time",
           subtitle = "The word 'protest' has always had a greater presence in media", x = "Year", y = "Apperence of Word", caption = "data: abcnews")  +
  theme(plot.title = element_text(face = "bold", size = 10),plot.subtitle = element_text(face = "italic"))
```
## Conclusion
We can see there is no clear trend in this graph. This allows us to assume that there has been no excessive use of either word by ABC.


## Presence of riot and protest in 2021
```{r plot-1: 2021}

data_abc <- read.csv("abcnews-date-text.csv")

data_abc |> 
  mutate(riot = str_detect(headline_text, "riot")) |> 
  mutate(protest = str_detect(headline_text, "protest")) |> 
  filter(protest == TRUE | riot == TRUE)  |> 
  mutate(protest = as.integer(protest), riot = as.integer(riot)) |> 
  mutate(year = substr(publish_date, 1, 4)) |> 
  mutate(month = substr(publish_date, 5, 6)) |> 
  group_by(month) |> 
  mutate(num_riot = sum(riot)) |> 
  mutate(num_protest = sum(protest)) |> 
  filter(year == "2021") |> 
  ggplot(aes(month)) +
      geom_point(aes(y = num_riot), color = "darkred") +
      geom_point(aes(y = num_protest), color = "blue") +
      theme_clean() +
   labs(title = "Presence of 'riot' and 'protest' in Australian Media During 2021",
           subtitle = "The word 'protest' still had a greater presence", x = "Month", y = "Apperence of Word", caption = "data: abcnews") +
  theme(plot.title = element_text(face = "bold", size = 10),plot.subtitle = element_text(face = "italic"))

```
## Conclusion
Even as we analyze the specific year we can see that there is no trend. Although this is not definitive proof that this source is reliable, it gives the reader a little more trust. 

## Source
My Source of Data was from ABC-Australia 
*the link to download the data can be found at <https://www.kaggle.com/datasets/therohk/million-headlines
