---
title: "Presence of Polarzing Words in Australian Media"
description: |
    Suvan Dommeti
site: distill::distill_website
---

## Introduction

In this project I hope to investigate the presence of the word "protest" and "riot" in Australian media. After searching the web I found an data source containing article headline, content and publish date from the Australian Broadcasting Corporation or ABC.

## Importance

Its important to see if a news source is disproportionately using words like riot in their publications because its an indicator that their publications have some sort of bias. The goal of this project is to develop a method of identifying the number of times a keyword is present as well as displaying it. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#Distill
library(distill)

# Display Libraries
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(tidyr)
library(dplyr)
library(rmarkdown)

#Verification
search()
```

```{r reading csv, include=FALSE}

data_abc <- read.csv("abcnews-date-text.csv")

#Viewing Data
glimpse(data_abc)
```


## Table

First lets take a look at some of the headlines in Australian News. Its always smart to take a quick glance at your data before you use it incase there are any glaring problems. Lets also add columns to tell us if the word Riot or Protest is present within the headline

```{r, table, layout="l-body-outset"}
data_abc <- read.csv("abcnews-date-text.csv")
  
table <- (sample_n(data_abc, 100))

table1 <- table |> 
  arrange(publish_date) |> 
  select(headline_text) |> 
  mutate(riot = str_detect(headline_text, "riot")) |> 
  mutate(protest = str_detect(headline_text, "protest")) 

paged_table(table1)
```

Ok now that we took a look at our data, lets start graphing. To graph Categorical Data we need to first quantify our variables. To do this lets count the number of occurrences of "True" within each column. Once we have a number
we can group by year.

Now lets graph!


## Presence of riot and protest from 2003 to 2021



```{r plot-1: years}

#plot 1 is in years and is displayed right here
data_abc <- read.csv("abcnews-date-text.csv")

colors <- c("Riot" = "tomato", "Protest" = "skyblue")
data_abc |> 
  mutate(riot = str_detect(headline_text, "riot")) |> 
  mutate(protest = str_detect(headline_text, "protest")) |> 
  filter(protest == TRUE | riot == TRUE)  |> 
  mutate(protest = as.integer(protest), riot = as.integer(riot)) |> 
  mutate(year = substr(publish_date, 1, 4)) |> 
    group_by(year) |> 
  mutate(num_riot = sum(riot)) |> 
  mutate(num_protest = sum(protest)) |> 
  ggplot(aes(year)) +
      geom_point(aes(y = num_riot, color = "Riot")) +
      geom_point(aes(y = num_protest, color = "Protest")) +
      theme_clean() +
      labs(title = "Presence of 'riot' and 'protest' in Australian Media Over Time",
           subtitle = "The word 'protest' has always had a greater presence in media", x = "Year", y = "Apperence of Word", caption = "data: abcnews", color = "Search Word" )  +
      theme(plot.title = element_text(face = "bold", size = 10),plot.subtitle = element_text(face = "italic"),
            axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = colors)
```

## Conclusion

We can see there is no clear trend in this graph. This allows us to assume that there has been no excessive use of either word by ABC.

## Presence of riot and protest in 2021

```{r plot-1: 2021}

data_abc <- read.csv("abcnews-date-text.csv")

colors <- c("Riot" = "tomato", "Protest" = "skyblue")

data_abc |> 
  mutate(riot = str_detect(headline_text, "riot")) |> 
  mutate(protest = str_detect(headline_text, "protest")) |> 
  filter(protest == TRUE | riot == TRUE)  |> 
  mutate(protest = as.integer(protest), riot = as.integer(riot)) |> 
  mutate(year = substr(publish_date, 1, 4)) |> 
  mutate(month = substr(publish_date, 5, 6)) |> 
  group_by(month) |> 
  mutate(num_riot = sum(riot)) |> 
  mutate(num_protest = sum(protest)) |> 
  filter(year == "2021") |> 
  ggplot(aes(month)) +
      geom_point(aes(y = num_riot, color = "Riot")) +
      geom_point(aes(y = num_protest, color = "Protest")) +
      theme_clean() +
      labs(title = "Presence of 'riot' and 'protest' in Australian Media During 2021",
            subtitle = "The word 'protest' still had a greater presence", 
             x = "Month", y = "Apperence of Word", caption = "data: abcnews",              color = "Search Word") +
      theme(plot.title = element_text(face = "bold", size = 10),
              plot.subtitle = element_text(face = "italic")) +
      scale_color_manual(values = colors)
```

## Conclusion

Even as we analyze the specific year we can see that there is no trend. Although this is not definitive proof that this source is reliable, it gives the reader a little more trust.

## Source

My data was collected from ABC-Australia by Rohit Kulkarni. \*the link to download the data can be found at \<https://www.kaggle.com/datasets/therohk/million-headlines>
